---
title: "EDAV Fall 2019 Problem Set 3"
author: "Swapnav Deka, Suman Tripathy"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

Read *Graphical Data Analysis with R*, Ch. 6, 7

Grading is based both on your graphs and verbal explanations. Follow all best practices as discussed in class.

Data cleaning: for many of the questions, you will have to clean up the data more than in past assignments. Labels do not have to perfect but they have to be legible. Often it is helpful to shorten or abbreviate labels: this can be done before plotting or at times within the plot functions. You may make any changes to the data as appropriate before plotting, including renaming column names and the like. Be sure though to include all adjustments in your scripts.

```{r setup, include=FALSE}
 # keep this chunk in your .Rmd file
 knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```
    
### 1. Coal Emissions

Data: https://datadiscovery.nlm.nih.gov/Environmental-Health/TOXMAP-EPA-Clean-Air-Markets-2016-Coal-Emissions/n97u-wtk7

(OK to manually download `.csv` file)

(a) Using `parcoords::parcoords()`, create an interactive parallel coordinate plot of `SO2`, `NOx`, `CO2`, `Heat Input` and `Gross Load`. Each line in the plot should represent one coal plant. Note any patterns.
```{r}
library(tidyverse)
library(parcoords)

coal <- read.csv('coal.csv')
coal %>% dplyr::select(5,7,9,11,12) %>% parcoords(rownames = F, alpha=0.4, reorderable = TRUE)

```

The majority of coal plants have SO2 levels between 0 and 5000 with some outliers (one in particular all the way above 40,000). For Gross Load, C02, and Heat Input, the lines tend to be very parallel, meaning there's a direct correlation between Heat Input and C02, CO2 and Gross Load, and Heat Input and Gross Load.


(b) Create a second interactive parallel coordiante plot, with the lines colored by state. Include only the states with more than 15 plants. Do any new patterns emerge? Explain briefly.
```{r}
library(d3r)

coal %>% 
  mutate(state = str_sub(Location,1,2)) %>% 
  add_count(state) %>%
  filter(n > 15) %>%
  dplyr::select(5,7,9,11,12,14) %>% 
  parcoords(rownames = F, color = list(colorBy = "state"),
            withD3 = TRUE, reorderable = TRUE, alpha = 0.4
            )

```

No major new patterns emerge.  We see that Texas has the two highest outliers for SO2 with one above 40,000 and another around 35,000.  Generally, we see that Indiana has high overall NOx levels.  We still see the direct correlations between Heat Input and CO2, CO2 and Gross Load, and Heat Input and Gross Load.   

(c) What do `SO2`, `NOx` and `CO2` stand for? Briefly describe why each is harmful.

SO2 is Sulfur Dioxide, and it's a sharp smelling gas that reacts easily with other chemicals to create harmful compounds. NOx is Nitric Oxide, a chemical compound that is combustible at high temperatures.  It's also an irritant gas that can inflame airways and prevent breathing at high concentrations. CO2 is carbon dioxide, a gas that causes the greenhouse effect by trapping heat within the earth's atmosphere. 


### 2. Planets

Using **rvest**, scrape the data from this table: https://nssdc.gsfc.nasa.gov/planetary/factsheet/ (hint: `html_table` is helpful). Remove `MOON` data. Then, using `GGally::ggparcoord()`, create two parallel coordinate plots of the numerical variables in the dataset, one colored by the value of `Ring System?` and one colored by planet name. In both cases, each line should represent one planet. Use `coord_flip` so the variable names are easier to read.  Describe any patterns you find.
```{r}
library(GGally)
library(rvest)

webpage <- read_html("https://nssdc.gsfc.nasa.gov/planetary/factsheet/")

planets <- webpage %>%
  html_node("table") %>%
  html_table(header=TRUE, fill=TRUE) %>%
  dplyr::select(names = 1, everything(),-5) %>%
  gather(variable, value, -names) %>% 
  spread(names, value) %>%
  dplyr::select(Planet=2, everything(), RingSystem=20,-c(8,22)) %>%
  mutate_all(funs(gsub(",", "", .))) %>%
  mutate_at(c(1,2,19), as.factor) %>%
  mutate_if(is.character, as.numeric) %>%
  column_to_rownames(var="variable")

ggparcoord(planets, c(2:17,19), groupColumn="RingSystem", scale='uniminmax') + coord_flip()
ggparcoord(planets, c(2:17,19), groupColumn="Planet", scale='uniminmax') + coord_flip()


```

Firstly, when exploring the dataframe, we found that 4 out of 9 surface pressure values are unknown, and the others vary a lot between 0 and 92, all to different levels of precision as well. Therefore, we have decided to drop this data column. We also drop all the categorical variables. Generally, we see that planets with a Ring System have a greater number of moons, higher gravity, higher escape velocity, and greater diameter. There's a lot of variation among many of the variables regardless of the ring system (i.e. we see low and high values for Aphelion and Perihelion for planets with and without a ring system).

```{r}
ggparcoord(planets, c(2:17,19), groupColumn="Planet", scale='uniminmax') + coord_flip()

```

Looking at the graph with data colored by PlanetName, we see that Saturn has the highest orbital period, orbital inclination, and orbital eccentricity. We see that Jupiter has the highest gravity and escape velocity. Pluto is furthest from the sun and has the largest aphelion value. All these observations make sense and match up to the dataframe.


### 3. Heart Disease

Data: four data frames that begin with `heart_disease` in the **ucidata** package

Packages: You may use **vcd** or **ggmosaic**.

(a) Create three mosaic plots, each involving two categorical variables from `heart_disease_cl` and interpret the plots. (You may reuse variables, for example X ~ Y and X ~ Z).
```{r}
library(ucidata)
library(vcd)

mosaic(cp ~ thal, 
       direction = c('v','h'),
       labeling = labeling_border(
         abbreviate_labs = c(3, 6),
         rot_labels=c(0,90)
         ),
       spacing = spacing_increase(),
       heart_disease_cl)

mosaic(slope ~ cp, 
       direction = c('v','h'),
       labeling = labeling_border(
         abbreviate_labs = c(5, 6),
         rot_labels=c(0,90)
         ),
       spacing = spacing_increase(),
       heart_disease_cl)

mosaic(cp ~ sex, 
       direction = c('v','h'),
       labeling = labeling_border(
         abbreviate_labs = c(6, 4),
         rot_labels=c(0,90)
         ),
       spacing = spacing_increase(),
       heart_disease_cl)

```

For the first plot, we see that most people have normal values for thal, some have reversable defects, and very few have fixed defects. For people with fixed and reversable defects, the majority of them have asympotatic chest pain. People with normal values for thal have the highest proportion of non-anginal pain.

For the second plot, we see that most people have asymptomatic chest pain, and the majority of people within that category have flat slopes of the peak exercise ST segment.  Upsloping slopes are most common for people with atypical angina chest pain and non-anginal chest pain. Of all types of chest pain, typical angina chest pain is the least common.  For the peak exercise ST segment, most people have either flat or upsloping slopes, and very few people see a downsloping trend.

For the third plot, we see that around 70% of our patients are male and 30% are female.  For males, the most common chest pain type is asymptomatic.  For females, there are similar proportions of non-anginal pain and asymptomatic pain.  For both genders, typical angina chest pain is the most rare, and atypical angina chest pain is the second most rare.


(b) Combine the four heart disease datasets and create a mosaic plot showing chest pain by sex and location. Describe any patterns.
```{r}
heart <- rbind(heart_disease_ch %>% mutate(location = 'Zurich'), 
               heart_disease_cl %>% mutate(location = 'Cleveland'), 
               heart_disease_hu %>% mutate(location = 'Hungary'), 
               heart_disease_va %>% mutate(location = 'Long Beach'))

mosaic(cp ~ location + sex , 
       heart, 
       labeling = labeling_border(
         abbreviate_labs = c(10, 1, 4),
         rot_labels=c(0,0,0)),
       spacing = spacing_increase(),
       direction = c('v','v','h')
       )

```

The most noticeable pattern is that Long Beach and Zurich both have extremely low proportions of females compared to males.  Cleveland and Hungary have the most patients.  In every division of gender and location, we see that asymptomatic chestpain is the most common type of chest pain, except for females in Hungary, where the most common type of chest pain is atypical angina pain.  In Cleveland and Hungary, we see that there's a lower proportion of asymptomatic chest pain in females as compared to males, whereas in Long Beach and Zurich, the proportions are relatively equal.  Also in every location, we see that there are fewer female patients than male.


### 4. District 3 Elementary Schools

Recently, there has been much debate about the lack of racial and economic diversity at Manhattan District 3 elementary schools, part of a larger and long-standing controversy about iniquities in the New York City public school system as a whole.

The *New York Times* article, "Rezoning Plan to Remake 3 Upper West Side Schools Will Proceed, City Says,"  (https://www.nytimes.com/2016/11/10/nyregion/rezoning-plan-for-3-upper-west-side-schools-will-proceed-city-says.html) (2016-11-10) identifies the 11 elementary schools in Manhattan District 3.

For this question, we will analyze parent survey results for these schools.

Data: https://www.schools.nyc.gov/about-us/reports/school-quality/nyc-school-survey


(a) Choose one of the likert style questions from the 2019 parent survey and use a diverging stacked bar chart to show results for the 11 schools identified in the article referenced above.
```{r}
library(readxl)
library(HH)

parent <- read_excel('parent.xlsx')

new <- parent %>% 
  filter(str_sub(DBN,1,3) == '03M') %>% 
  filter(str_sub(DBN,4,6) %in% c('009','165','145','163', '075', '084', '166', '087', '452', '199', '191')) %>%
  mutate(school = str_sub(DBN,4,6)) %>%
  dplyr::select('school'= school, 'SD' = 4, 'D' = 5, 'A' = 6, 'SA' = 7) %>%
  mutate_at(1, as.factor) %>%
  mutate_if(is.character, as.integer)

likert(school~., new, as.percent=FALSE, rightAxis = FALSE, main='"The school regularly communicates with me about my child." (2019)', xlab = 'percent')
```

(b) Choose a question that was asked in 2014 and 2019 and compare results for the three schools discussed most in the article: P.S. 199, P.S. 191, and P.S. 452. You may use two separate diverging stacked bar charts or combine all the information in one.
```{r}
parent2014 <- read_excel('2014parent.xlsx')

new2019 <- parent %>% 
  filter(str_sub(DBN,1,3) == '03M') %>% 
  filter(str_sub(DBN,4,6) %in% c('452', '199', '191')) %>%
  mutate(school = str_sub(DBN,4,6)) %>%
  dplyr::select('school'= school, 'SD' = 77, 'D' = 78, 'A' = 79, 'SA' = 80) %>%
  mutate_at(1, as.factor) %>%
  mutate_if(is.character, as.integer)

new2014 <- parent2014 %>% 
  filter(str_sub(DBN,1,3) == '03M') %>% 
  filter(str_sub(DBN,4,6) %in% c('452', '199', '191')) %>%
  mutate(school = str_sub(DBN,4,6)) %>%
  dplyr::select('school'= school, 'SD' = 67, 'D' = 66, 'A' = 65, 'SA' = 64) %>%
  mutate_at(1, as.factor) %>%
  mutate_if(is.character, as.integer)


likert(school~., new2014, as.percent=FALSE, rightAxis = FALSE, main='"I feel that my child is safe at school." (2014)', xlab = 'percent')
likert(school~., new2019, as.percent=FALSE, rightAxis = FALSE, main='"I feel that my child is safe at school." (2019)', xlab = 'percent')


```

(c) Interpret your findings of (b) in light of the reputations of the schools as presented in the article. Are they surprising or what you would have expected?

The 2014 reponses come from data prior to the school district voting on the resolutions to redistrict the schools in the Upper West Side. School 452 and 199 were identified as schools with predominantly well-off, white population. Many parents in these schools' boundaries are the ones who threatened lawsuits in response to possible redistricting, so it is reasoned that this population would be comfortable with the current situation. This is mirrored in the dataset, showing that 100% of respondents from school 452 either agreed or strongly agreed that they felt their children were safe at school, and 99% of parents from School 199 felt the same way. School 191 was identified as being in a poorer minority neighborhood, but parents from this area still felt relatively safe, at a rate of 96%. However, following the redistricting, these valuations changed significantly. School 452 now has the highest rate of unsafe responses. This makes sense because the school was moved to the location of a school that parents from 452 originally felt was unsafe, so they are likely to feel relatively unsafe compared to the previous location. It also makes sense that the population of 452 parents whose children were moved to 191 and 199 would also feel relatively unsafe compared to their old situation. For this reason, it makes sense and is supported by the data that all three schools have a higher proportion of parents who feel relatively unsafer than they did prior to redistricting.


